Building and Checking Suffix Array Using Induced Sorting Method

### 2017-08-05 new-draft/draft.tex
2017/08/06

Suffix array~(SA) can be built by the induced sorting~(IS) method on both internal and external memory models. We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously. Particularly, the first method checks both suffix and longest common prefix~(LCP) arrays for an input string drawn from a full-ordered alphabet $\Sigma$. By combining our methods with the Karp-Rabin fingerprinting function, we design two probabilistic algorithms that perform verification with a negligible error probability. Theoretically, the algorithm designed by the first method has a sorting complexity, while the algorithm designed by the second method takes linear time and $\mathcal{O}(\|\Sigma\|)$ RAM space. We integrate the algorithm designed by the second method into the existing suffix sorting algorithm DSA-IS for performance evaluation. Our experiments indicate that the time, space and I/O consumptions for verifcation are considerably smaller than that for construction. We believe that the proposed methods, together with the state-of-the-art IS suffix sorters (and LCP-array builders), could constitute efficient solutions for building and checking suffix (and LCP) array(s) when $\|\Sigma\|$ is much smaller than the size of input string.

@@

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Suffix and LCP arrays, construction and verification, external memory.
\end{IEEEkeywords}}

@@

A constructed SA should be checked to detect potential errors caused by implementation bugs and hardware malfunctions. The software packages for DC3 and eSAIS provide users a checker that implements the algorithm presented in~\cite{Dementiev2008a}. When running on external-memory model, the checking overhead is rather high because this checker takes two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples. In this paper, we describe two methods that enable any IS suffix sorter to build and check an SA at the same time. In particular, the first method is proposed for checking both suffix and LCP arrays. We apply the Karp-Rabin fingerprinting function~\cite{Karp1987} to designing two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability. We combine the algorithm designed by the second method with DSA-IS and implement the combination to make a performance evaluation, where new substring sorting and naming methods is used to lower the peak disk use and I/O volume of the program for DSA-IS in advance. Our experimental results indicate that the time, space and I/O consumptions for verification are negligible to that for construction. This implies that the IS method can be used to design efficient solutions for verification as well.

@@

Partition $x$ into multiple successive S*-type substrings, such that substring starts with an S*-type character and ends with the leftmost S*-type character on the right side. Produce a reduced string $x_1$ by replacing each S*-type substring with its name, where the name indicats its relative order among all.

@@

Algorithm~\ref{alg:1} shows the framework of an IS suffix sorting algorithm. At the very beginning, a reduction phase for sorting and naming S*-type substrings is called to produce the reduced string $x_1$. If there exist duplicate characters in $x_1$, the reduction phase is recursively called with $x_1$ as input; otherwise, all the S*-type suffixes are already sorted and $sa_1$ is directly computed from $x_1$. 
Afterward, an induction phase for sorting all the suffixes is called to produce $sa$ for the input string at current level. The induction phase is recursively called with $sa$ as input until reaching the top recursion level. During the execution of a reduction/induction phase, the order of substrings/suffixes is determined by comparing their heading characters and the ranks of their successors according to the IS principle, which involves a great number of random accesses to $x$ and $sa$. This can be done very fast if both $x$ and $sa$ wholly reside on RAM; otherwise, each access takes an individual I/O operation, leading to a severe performance degradation. This problem is solved in DSA-IS by performing two steps below during the reduction/induction phase:

@@

After the first two steps, the short S*-type substrings of each block and the long S*-type substrings of $x$ are sorted and separately organized as a sequence in external memory. Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieve the top item from the heap to determine the lexical order of substrings from different sequences. The heap contains at most $k + 1$ substrings at any point of time and it compares any two substrings in $\mathcal{O}(D)$ time by a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}. Theoretically, the sorting method has a good performance if the majority of S*-type substrings in $x$ are short with a small $D$. This is commonly satisfied in practice, because the average length of S*-type substrings is typically small in real-world datasets.

In what follows, we describe a method for naming the S*-type substrings during the above sorting process. The key point here is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}' as fast as possible. If either of the two substrings is short, our solution is to literally compare them in $\mathcal{O}(D)$ time. Otherwise, if both of them are long, then we assign names to all the long S*-type substrings using the technique proposed for SAIS-PQ in S\ref{dsaism_sorting_method:2}' and compare two successively popped long S*-type substrings by their names. This technique is based on the idea that two substrings are of the same name if their heading characters and the names of their successors are both equal\footnote{We refer interested readers to~\cite{Liu15} for more information.}.

@@

The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $n$ fixed-size tuples using external memory. As can be observed from Section~\ref{sec:experiments}, the peak disk use and the I/O volume for an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.

@@

Recall that, an IS suffix sorting algorithm recursively perform the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters. Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase mainly consists of the following steps:

@@

A running example of S\ref{induction_phase:2}-S\ref{induction_phase:4} is shown in Fig.~\ref{fig:example1}. Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$. For example, $\sf suf(8)$, $\sf suf(5)$, $\sf suf(2)$ are placed into the right part of $\sf sa\_bkt(i)$ and arranged in their sorted order indicated by $sa^*$. Then, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes. To do this, we first check $sa[0] = 14$ (marked by the symbol @) and induce the predecessor of $\sf suf(14)$ in lines 8-9. Because $x[13] = i$ is L-type, we put $\sf suf(13)$ into the current leftmost empty position in $\sf sa\_bkt_L(i)$. To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17. Afterward, we find the rightmost position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes. When scanning $sa[14] = 4$ in lines 19-20, we see $x[3] = i$ is S-type and thus put $\sf suf(3)$ into the current rightmost empty position in $\sf sa\_bkt_S(i)$. Following the same idea, we get all the S-type suffixes sorted in line 28. Based on the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient and necessary conditions for SA verification.

@@

\begin{lemma} \label{lemma:2}
    $sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied: \\
    (1) $sa^*$ is correctly computed. \\
    (2) S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented. \\
\end{lemma}

\begin{IEEEproof} Given two suffixes placed at $sa[i]$ and $sa[j]$ and their successors placed at $sa[p]$ and $sa[q]$, we prove the statement as follows.

Because S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented, $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$. This satisfies the second condition of Lemma~\ref{lemma:1}.

Further, suppose $sa[i] = sa[j]$ and $i \ne j$, then $sa[p] = sa[q]$ and $p \ne q$. Repeatedly replacing $(i, j)$ with $(p, q)$ until $\sf suf(sa[i])$ is S*-type, then $sa^*$ contain duplicate elements. Because $sa^*$ is correctly computed, each element in $sa^*$ must be unique. This leads to a contradiction.
\end{IEEEproof}

The above lemma indicates that, if the first condition of Lemma~\ref{lemma:2} is true, we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification. It should be noticed that the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines. This is considerable smaller than the whole program for a disk-based suffix sorting algorithm. Following the idea, we assume S\ref{induction_phase:1}-S\ref{induction_phase:4} are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and I/O errors caused by hardware malfunctions. The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA at the same time.

@@

Our first method is directly extended from Lemma~\ref{lemma:2}.

\begin{lemma} \label{lemma:3}
	Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\
	(1) $sa^*$ is correct. \\
	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\

@@

The first condition of Lemma~\ref{lemma:3} can be checked by a sparse SA checker like~\cite{wu2017}. Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:2}'' and S\ref{induction_phase:4}''. The second condition intends for detecting I/O errors and can be checked by determining whether the induced and scanned values for each suffix are equal. The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once. Our solution is to integrate the checking process into the building process, by ensuring the sequence of values placed at a bucket is identical to that scanned later. For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end of the induction phase. If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability. As a result, $sa$ can be built and probabilistically checked at the same time. It should be noticed that this method can be also applied to checking the longest common prefix array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}. As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ during the induction phase at the top recursion level.

@@

	%\SetAlgoNoLine
	\Fn{{\sf CheckByMethodA}($x, sa^*, lcp^*$))}{

        Check the correctness of $sa^*$ and $lcp^*$ by the method proposed in~\cite{wu2017}. \\

@@

Method B uses a different way to check the correctness of $sa^*$ in Lemma~\ref{lemma:4}. Before our presentation, we first introduce a notation called $\overline{sa^*}$. The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes. Both of them are computed during the induction phase at the top recursion level of an IS suffix sorting algorithm. To be specific, $sa^*$ is computed from $sa_1$ and $\overline{sa^*}$ is retrieved from the constructed $sa$.

\begin{lemma} \label{lemma:4}
    Assume the induction phase at the top recursion level is correctly implemented, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied for all $i \in [0, n_1)$:

	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes. \\
	(2) $sa^*[i] = \overline{sa^*}[i]$. \\
	(3) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}''. \\

@@

The proposed two methods check equality of two arrays by comparing their fingerprints, which can be calculated using a hash function. Notice that two equal arrays must share an identical hash value, but the inverse is not always true. To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute the required values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $delta$ is an integer randomly chosen from $[1, L)$. By setting $L$ to a large value, the error probability can be reduced to a negligible level. In Fig.~\ref{fig:example2}, we depict an example for computing the integer array $A$, where $A$ is identical to $sa^*$ and $\overline{sa^*}$ in Fig.~\ref{fig:example1}. Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4. Because $\sf fp(A[0, -1]) = 0$, $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$. By iteratively computing the fingerprints of the prefixes of $A$ according to Formula~\ref{formula:2}, we finally obtain the fingerprints of the whole array in line 7.

@@

It 
According to Lemma~\ref{lemma:3}, Method A must check both $sa^*$ and $sa$ to accomplish verification. Similar to Method B, the overhead for checking $sa$ is mainly caused by fingerprint calculations and thus can be neglected. On the other hand, we can apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within sorting complexity, where the time and space in need is proportional to the number of S*-type characters in $x$. Because the ratio of S*-type characters to all in $x$ is commonly one-third in real-world datasets, the checking process for $sa^*$ will not become the bottleneck for Method A. To achieve a higher speed, we can also parallelize the checking processes for $sa$ and $sa^*$ using more computing resources.

Assuming the induction phase at the top recursion level is correctly implemented, the bottleneck of Algorithm~\ref{alg:2} and~\ref{alg:3} lie in checking the correctness of $sa^*$. For Algorithm~\ref{alg:2}, we apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within the sorting complexity. Specifically, given two suffixes, i.e. $\sf suf(i)$ and $\sf suf(j)$, and their LCP-value $d$, we compute and compare the fingerprints of $\sf sub(i, i + d - 1)$ and $\sf sub(j, j + d - 1)$. If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ corresponds to that of $\sf suf(i)$ and $\sf suf(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability. Because at most one out of every two suffixes is S*-type (commonly one-third in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption. For Algorithm~\ref{alg:3}, the checking overhead is mainly caused by fingerprinting calculations and thus can be neglected compared with the building consumption.

@@

Rather than designing an I/O layer for efficient I/O operations, we currently use the containers provided by the STXXL library to perform reading, writing and sorting in external memory, these containers do not free the disk space for storing temporary data even if it is not needed any more, leading to a space consumption higher than our expectation. This is an implementation issue that can be solved by storing the data into multiple files and deleting each file when it is obsolete. In this way, our program can be further improved to achieve a space performance comparable to fSAIS.

@@

In this paper, we proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously. In particular, the probabilistic algorithm designed by the second method is rather lightweight in terms of that it takes negligible time and space compared with the existing IS suffix sorting and checking algorithms. We also made an attempt to improve the space performance of DSA-IS using new substring sorting and naming methods. The experimental results shows that our program for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and requires only half as much disk space as the latter on various real-world datasets. We also showed that the space performance of this program can be further optimized using some optimization techniques. In our next paper, we will propose another disk-based suffix sorting algorithm that takes only $1n$ work space to run.

>>>>

Suffix array~(SA) can be built by the induced sorting~(IS) method on both internal and external memory models. We propose two methods that enable any IS suffix sorter to build and check an SA simultaneously. Particularly, the first method checks both suffix and longest common prefix~(LCP) arrays for an input string drawn from a full-ordered alphabet $\Sigma$. By combining our methods with the Karp-Rabin fingerprinting function, we design two algorithms that perform verification with a negligible error probability. Theoretically, the algorithm designed by the first method has a sorting complexity, while the algorithm designed by the second method takes linear time and $\mathcal{O}(\|\Sigma\|)$ RAM space. We integrate the algorithm designed by the second method into the existing suffix sorting algorithm DSA-IS for performance evaluation. Our experiments indicate that the time, space and I/O consumptions for verifcation are considerably smaller than that for construction. We believe that the proposed methods, together with the state-of-the-art IS suffix sorters (and LCP-array builders), could constitute efficient solutions for building and checking suffix (and LCP) array(s) when $\|\Sigma\|$ is much smaller than the size of input string.

@@

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Suffix and LCP arrays, construction and verification, internal and external memory.
\end{IEEEkeywords}}

@@

A constructed SA should be checked to detect potential errors caused by implementation bugs and hardware malfunctions. The software packages for DC3 and eSAIS provide users a checker that implements the algorithm presented in~\cite{Dementiev2008a}. When running on external-memory model, the checking overhead is rather high because this checker takes two passes of integer sorts for arranging $\mathcal{O}(n)$ fixed-size tuples. In this paper, we describe two methods that enable any IS suffix sorter to build and check an SA at the same time. In particular, the first method is proposed for checking both suffix and LCP arrays. We apply the Karp-Rabin fingerprinting function~\cite{Karp1987} to designing two probabilistic algorithms, in terms of that their checking results are wrong with a negligible probability. We combine the algorithm designed by the second method into DSA-IS and implement the combination to make a performance evaluation, where new substring sorting and naming methods are used to redesign DSA-IS for lowering the peak disk use and I/O volume taken by SA construction. Our experimental results indicate that the time, space and I/O consumptions for verification are negligible to that for construction. This implies that the IS method can be used to design efficient solutions for verification as well.

@@

Partition $x$ into multiple successive S*-type substrings, such that eachsubstring starts with an S*-type character and ends with the leftmost S*-type character on the right side. Produce a reduced string $x_1$ by replacing each S*-type substring with its name, where the name is the rank indicating the lexical order among all.

@@

Algorithm~\ref{alg:1} shows the framework of an IS suffix sorting algorithm. At the very beginning, a reduction phase for sorting and naming S*-type substrings is called to produce the reduced string $x_1$. If there exist duplicate characters in $x_1$, the reduction phase is recursively called with $x_1$ as input; otherwise, all the S*-type suffixes are already sorted and $sa_1$ is directly computed from $x_1$.
Afterward, an induction phase for sorting all the suffixes is called to produce $sa$ for $x$ at current recursion level. The induction phase is recursively called with $sa$ as input until reaching the top recursion level. During the execution of a reduction/induction phase, the order of substrings/suffixes is determined by comparing their heading characters and the ranks of their successors according to the IS principle, which involves a great number of random accesses to $x$ and $sa$. This can be done very fast if both $x$ and $sa$ wholly reside on RAM; otherwise, each access takes an individual I/O operation, leading to a severe performance degradation. This problem is solved in DSA-IS by performing two steps below during the reduction/induction phase:

@@

After the first two steps, the short S*-type substrings of each block and the long S*-type substrings of $x$ are sorted and separately organized as a sequence in external memory. Assume $x$ is split into $k$ blocks, the multi-way sorter in S\ref{dsaism_sorting_method:3}' maintains an internal-memory heap to cache the current smallest of each sequence and continually retrieves the top item from the heap to determine the lexical order of substrings from different sequences. The heap contains at most $k + 1$ substrings at any point of time and it compares any two substrings in $\mathcal{O}(D)$ time by a literal string comparison\footnote{If the leftmost $D$ characters of a long S*-type substring is equal to a short S*-type substring, then the short is lexicographically greater than the long.}. Theoretically, the sorting method has a good performance if the majority of S*-type substrings in $x$ are short with a small $D$. This is commonly satisfied in practice, because the average length of S*-type substrings is typically small in real-world datasets.

In what follows, we describe a method for naming the S*-type substrings during the above sorting process. The key point here is to check equality of two substrings successively popped from the heap in S\ref{dsaism_sorting_method:3}' as fast as possible. If either of the two substrings is short, we literally compare them in $\mathcal{O}(D)$ time. Otherwise, if both of them are long, then we assign names to all the long S*-type substrings using the technique proposed for SAIS-PQ in S\ref{dsaism_sorting_method:2}' and directly compare two successively popped long S*-type substrings by their names. This technique is based on the idea that two substrings are of the same name if their heading characters and the names of their successors are both equal\footnote{We refer interested readers to~\cite{Liu15} for more information.}.

@@

The disk-based implementation of this checker conducts two passes of integer sorts and each sort arranges the order of $\mathcal{O}(n)$ fixed-size tuples using external memory. As can be observed from Section~\ref{sec:experiments}, the peak disk use and the I/O volume for an SA encoded by 40-bit integers are around $26n$ and $53n$, respectively.

@@

Remember that an IS suffix sorting algorithm recursively performs the reduction phase to compute the reduced string $s_1$ until $s_1$ contains no duplicate characters. Afterward, it produces $sa_1$ from $s_1$ and recursively perform the induction phase to compute $sa$ until reaching the top recursion level, where the induction phase consists of the following steps:

@@

A running example of S\ref{induction_phase:2}-S\ref{induction_phase:4} is shown in Fig.~\ref{fig:example1}. Given $sa^*$ is already known, line 6 inserts each S*-type suffix into the right part of the corresponding bucket and arranges them according to their sorted order indicated by $sa^*$. For example, $\sf suf(2)$, $\sf suf(5)$, $\sf suf(8)$ are placed at the rightmost empty position of $\sf sa\_bkt(i)$ in sequence. Then, we find the leftmost position of each bucket (marked by the symbol $\wedge$) and scan $sa$ rightward for inducing the order of L-type suffixes. To do this, we first check $sa[0] = 14$ (marked by the symbol @) and induce the predecessor of $\sf suf(14)$ in lines 8-9. Because $x[13] = i$ is L-type, we put $\sf suf(13)$ into the current leftmost empty position in $\sf sa\_bkt_L(i)$. To step through $sa$ in this way, we get all the L-type suffixes sorted in line 17. Afterward, we find the rightmost position of each bucket and scan $sa$ leftward for inducing the order of S-type suffixes. When scanning $sa[14] = 4$ in lines 19-20, we see $x[3] = i$ is S-type and thus put $\sf suf(3)$ into the current rightmost empty position in $\sf sa\_bkt_S(i)$. Following the same idea, we get all the S-type suffixes sorted in line 28. Based on the discussion, we show in Lemma~\ref{lemma:2} a set of sufficient and necessary conditions for SA verification.

@@

\begin{lemma} \label{lemma:2}
    $sa[0, n)$ is the SA for $x[0, n)$ if and only if the following conditions are satisfied: \\
    (1) $sa^*$ is correctly computed. \\
    (2) S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented at the top recursion level. \\
\end{lemma}

\begin{IEEEproof} Given two suffixes placed at $sa[i]$ and $sa[j]$ and their successors placed at $sa[p]$ and $sa[q]$ at the top recursion level, we prove the statement as follows.

Because S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented, we have $i < j \iff (x[sa[i]], p) < (x[sa[j]], q)$ following the IS principle. This satisfies the second condition of Lemma~\ref{lemma:1}.

Further, suppose $sa[i] = sa[j]$ and $i \ne j$, then $sa[p] = sa[q]$ and $p \ne q$. Repeating this reasoning process by replacing $(i, j)$ with $(p, q)$ until $\sf suf(sa[i])$ is S*-type, then $sa^*$ contains duplicate elements. Because $sa^*$ is correctly computed, each element in $sa^*$ must be unique. This leads to a contradiction.
\end{IEEEproof}

The above lemma indicates that, if the first condition of Lemma~\ref{lemma:2} is true, we can check the correctness of $sa^*$ instead of $sa$ at the top recursion level to perform SA verification. It should be noticed that the code snippet for the induction phase at the top recursion level only consists of tens of C++ code lines. This is considerable smaller than the whole program for a disk-based suffix sorting algorithm. Following the idea, we assume S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' are correctly implemented and propose two methods for checking computation errors caused by implementation bugs and hardware malfunctions. The algorithms designed by these methods can be seamlessly integrated into any IS suffix sorting algorithms for building and checking an SA at the same time.

@@

Method A is based on Lemma~\ref{lemma:3}, which is extended from Lemma~\ref{lemma:2}.

\begin{lemma} \label{lemma:3}
	Assume the induction phase is correctly implemented at the top recursion level, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied: \\
	(1) $sa^*$ is correct. \\
	(2) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' for $i \in [0, n)$. \\

@@

The first condition of Lemma~\ref{lemma:3} can be checked by a sparse SA checker like~\cite{wu2017}. Notice that each suffix induced into $sa$ will be latter scanned for inducing the order of its predecessor during S\ref{induction_phase:3}''-S\ref{induction_phase:4}''. The second condition intends for detecting errors caused by hardware malfunctions and can be checked by determining whether the induced and scanned values for each suffix are equal. The problem to be solved here is that when a suffix is induced into a bucket, its corresponding value in $sa$ may not be scanned at once. Our solution is to integrate the checking process into the building process, by ensuring the sequence of values placed at a bucket is identical to that scanned later. For the purpose, we use a fingerprinting function to increasingly compute the fingerprints of both sequences and check their equality in constant time at the end. If the two fingerprints for each bucket are equal, then the second condition of Lemma~\ref{lemma:3} will be seen with a high probability. As a result, $sa$ can be built and probabilistically checked at the same time. It should be noticed that this method can be also applied to checking the LCP array when it is being built from $lcp^*$ following the IS principle~\cite{Fischer11}.

As a result, we design Algorithm~\ref{alg:2} to probabilistically check $sa$ and $lcp$ during the induction phase at the top recursion level. We point out that ${\sf sa_{I1}}(c)$ and ${\sf sa_{I2}}(c)$ are two sequences respectively induced into ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$, while ${\sf sa_{S1}}(c)$ and ${\sf sa_{S2}}(c)$ are two sequences respectively scanned from ${\sf sa\_bkt_L}$ and ${\sf sa\_bkt_S}$.

@@

	%\SetAlgoNoLine
	\Fn{{\sf CheckByMethodA}($x, sa^*, lcp^*$)}{

        Verify $sa^*$ and $lcp^*$ by the checker presented in~\cite{wu2017}. \\

@@

As can be seen in Lemma~\ref{lemma:4}, Method B uses a different way to check the correctness of $sa^*$. Before our presentation, we first introduce a notation called $\overline{sa^*}$. The same as $sa^*$, $\overline{sa^*}$ also represents the suffix array for all the S*-type suffixes. Both of them are computed during the induction phase at the top recursion level. To be specific, $sa^*$ is computed from $sa_1$ and $\overline{sa^*}$ is retrieved from the constructed final $sa$.

\begin{lemma} \label{lemma:4}
    Assume the induction phase is correctly implemented at the top recursion level, the output $sa[0, n)$ of an IS suffix sorting algorithm is the SA for the input string $x[0, n)$ if and only if the following conditions are satisfied:

	(1) $sa^*[i]$ is a permutation of all the S*-type suffixes. \\
	(2) $sa^*[i] = \overline{sa^*}[i]$ for $i \in [0, n_1)$. \\
	(3) $sa[i]$ is equal to the value calculated by S\ref{induction_phase:1}''-S\ref{induction_phase:4}'' for $i \in [0, n)$. \\

@@

The proposed two methods check equality of two arrays by comparing their fingerprints, which can be calculated using a hash function. Notice that two equal arrays must share an identical hash value, but the inverse is not always true. To lower the error probability of a false match, we prefer using the Karp-Rabin fingerprinting function to compute the required values following Formulas~\ref{formula:1}-\ref{formula:2}, where $L$ is a prime and $\delta$ is an integer randomly chosen from $[1, L)$. By setting $L$ to a large value, the error probability can be reduced to a negligible level. In Fig.~\ref{fig:example2}, we depict an example for computing the integer array $A$, where $A$ is identical to $sa^*$ and $\overline{sa^*}$ in Fig.~\ref{fig:example1}. Given $L = 23$ and $\delta = 5$, we first compute the fingerprint of $A[0, 0]$ in line 4. Because $\sf fp(A[0, -1]) = 0$, then $\sf fp(A[0, 0]) = (0 \cdot 5 + 14) \, mod \, 23 = 14$. By iteratively computing the fingerprints of the prefixes of $A$ according to Formula~\ref{formula:2}, we finally obtain the fingerprints of the whole array in line 7.

@@

The bottleneck of Algorithm~\ref{alg:2} and~\ref{alg:3} lie in checking the correctness of $sa^*$. For Algorithm~\ref{alg:2}, we apply the method proposed in~\cite{wu2017} to ensure the correctness of $sa^*$ within the sorting complexity. Specifically, given two suffixes, i.e. $\sf suf(i)$ and $\sf suf(j)$, and their LCP-value $d$, we compute and compare the fingerprints of $\sf sub(i, i + d - 1)$ and $\sf sub(j, j + d - 1)$. If the fingerprints are equal and the order of $x[i + d]$ and $x[j + d]$ corresponds to that of $\sf suf(i)$ and $\sf suf(j)$, then the suffixes are correctly sorted and their LCP-value is right with a high probability. Because at most one out of every two suffixes is S*-type (commonly one-third suffixes are S*-type in real-world data sets) and the checking process is only executed during the induction phase at the top recursion level, the checking overhead is much less than the building consumption. For Algorithm~\ref{alg:3}, the checking overhead is mainly caused by fingerprinting calculations and thus can be neglected compared with the building consumption.

@@

Rather than designing an I/O layer for efficient I/O operations, we currently use the containers provided by the STXXL library to perform reading, writing and sorting in external memory, these containers do not free the disk space for storing temporary data even if it is not needed any more, leading to a space consumption higher than our expectation. This is an implementation issue that can be solved by storing the data into multiple files and deleting each file when it is obsolete. In this way, our program can be further improved to achieve a space performance comparable to fSAIS. In our next paper, we will propose another disk-based sorting algorithm that has a space performance even better than fSAIS. It takes only $1n$ work space to run excluding the input and output.

@@

We proposed two methods that enable any IS suffix sorting algorithm to build and check SA simultaneously. The program for our algorithm designed by Method B consumes negligible time and RAM space in comparison with that for the existing suffix sorting and checking algorithms on various real-world data sets. We also made our first attempt to improve the performance of DSA-IS using new substring and naming methods. The implementation for the adapted algorithm DSA-IS+ runs as fast as that for eSAIS and requires only half as much disk space as the latter. These results indicate the great potential of the IS method to design efficient solutions for SA constructing and verification.
